---
title: "IS2000 Assignment / Getting to Know R"
author: "Carminati, Heather"
date: "Spring 2024"
output:
  html_document:
    df_print: paged
---

## Question 2

```{r Q1_LoadCSV, echo = FALSE} 
# Load the CSV file into a data frame named df.flights and prints the first 6 rows of the data frame

df.flights <- read.csv(url("https://s3.us-east-2.amazonaws.com/artificium.us/datasets/FlightsWithAirlines.csv"),
                       stringsAsFactors = FALSE)

print(head(df.flights))
```

## Question 3

The dataset contains **`r nrow(df.flights)`** observations (rows) and **`r ncol(df.flights)`** columns. Each row is information about a flight, including departure and departure delay information. The average mean departure delay is **`r round(mean(df.flights$dep_delay))`** minutes.  A total of **`r sum(df.flights$dep_delay > 60)`** flights experienced a delay of more than an hour.

## Question 4

```{r Q4_IsNight, echo = FALSE} 
# Creates a new column in the data set, minutesWithDelay, that represents the total minutes with delay. 
df.flights$minutesWithDelay <- df.flights$dep_min + df.flights$dep_delay

# Creates a new column in the data set, hoursDelayed, that represents the concerted total minutes with delay to hours
df.flights$hoursDelayed <- df.flights$minutesWithDelay %/% 60

# Creates a new column in the dataset, isNight, that represents whether the flight is at night. 
df.flights$isNight <- ifelse(
  ((df.flights$dep_hr + df.flights$hoursDelayed) %% 24) > 19 | 
    ((df.flights$dep_hr + df.flights$hoursDelayed) %% 24) < 5, 
  TRUE, FALSE)

# Removes the temporary columns.
df.flights$minutesWithDelay <- NULL
df.flights$hoursDelayed <- NULL

# Prints the first two rows of the update data frame and prints the isNight column values.  
print(head(df.flights, 2))
print(df.flights$isNight)
```
